---
title: "Investigate-patients-with-ovarian-samples"
author: "Christelle Colin-Leitzinger"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: paged
    code_folding: hide
  word_document:
    toc: true
editor_options:
  chunk_output_type: console
---

<style type= "text/css">

. figure {
   margin-top: 25px;
   margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```

```{r library}
library(here)
library(tidyverse)
library(gtsummary)
library(labelled)
library(survival)
library(survminer)
theme_set(theme_classic())
theme_gtsummary_compact()
```

```{r load}
gyn_data <-
  read_rds(paste0(here(), 
                  "/data/processed data",
                  "/ORIEN_data_with_ovarian_sample2025-05-07.rds")) %>% 
  janitor::clean_names()
```

We have `r nrow(gyn_data)` patients with germline and tumor samples coolected in "GYN - Ovarian Cancer" disease type.

# Disease and histology
These are characteristics listed in the sequencing sample data.    
Some samples are not collected in the ovary. I talked with Yi-Han then Nancy. Yi-Han updated her samples selection filters to include ovary samples as first choice when available.
```{r}
gyn_data %>% 
  select(disease_type, tumor_specimen_site_of_collection, histology_behavior
         ) %>% 
  tbl_summary(sort = list(everything() ~ "frequency")) %>%
  bold_labels() %>%
  add_stat_label() %>%
  modify_header(label ~ "**Disease characteristics**",
                all_stat_cols() ~ "**{level}**, N = {n}")
```
<br>

# Let's clean primary cancer first
## Cancer characteristics before primary cancer filtering
You can see that 18 patients had a primary diagnosis of breast cancer and 12 of them had a subsequent diagnosis of ovarian cancer.    
I can extract the histology of the subsequent diagnosis of ovarian cancer if we want to include them.
```{r}
gyn_data %>% 
  # mutate(primary_diagnosis_site = relevel(factor(primary_diagnosis_site), ref = "Ovary")) %>% 
  mutate(primary_diagnosis_site = forcats::fct_infreq(primary_diagnosis_site)) %>% 
  # mutate(primary_diagnosis_site = relevel(factor(primary_diagnosis_site), ref = "Ovary")) %>% 
  select(disease_type, 
         histology, 
         primary_diagnosis_site, 
         
         has_subsequent_cancer, subsequent_cancer_primary_diagnosis_site, 
         
         primary_diagnosis_site_code,
         clin_group_stage, other_staging_system, other_staging_value, 
         ecog, karnofsky) %>% 
  tbl_summary(by = primary_diagnosis_site,
              sort = list(everything() ~ "frequency")) %>%
  bold_labels() %>%
  add_stat_label() %>%
  modify_header(label ~ "**Cancer characteristics**",
                all_stat_cols() ~ "**{level}**, N = {n}") %>% 
  modify_spanning_header(everything() ~ "Primary Diagnosis Site")
```
<br>

```{r clean primary site, echo = TRUE, class.source= 'fold-hide'}
gyn_data <- gyn_data %>%
  filter(primary_diagnosis_site %in% c(
    "Ovary", "Fallopian tube", "Peritoneum, NOS", "Specified parts of peritoneum"
  ))
```

## Cancer characteristics after primary cancer filtering

We now have `r nrow(gyn_data)` patients after filtering for Primary diagnosis site.

```{r}
gyn_data %>% 
  select(disease_type, tumor_specimen_site_of_collection#, histology_behavior
         ) %>% 
  tbl_summary(sort = list(everything() ~ "frequency")) %>%
  bold_labels() %>%
  add_stat_label() %>%
  modify_header(label ~ "**Disease characteristics**",
                all_stat_cols() ~ "**{level}**, N = {n}")
```
<br>

```{r}
gyn_data %>% 
  # mutate(primary_diagnosis_site = relevel(factor(primary_diagnosis_site), ref = "Ovary")) %>% 
  mutate(primary_diagnosis_site = forcats::fct_infreq(primary_diagnosis_site)) %>% 
  # mutate(primary_diagnosis_site = relevel(factor(primary_diagnosis_site), ref = "Ovary")) %>% 
  select(disease_type, 
         histology, 
         primary_diagnosis_site, 
         
         has_subsequent_cancer, subsequent_cancer_primary_diagnosis_site, 
         
         primary_diagnosis_site_code,
         clin_group_stage, other_staging_system, other_staging_value, 
         ecog, karnofsky) %>% 
  tbl_summary(by = primary_diagnosis_site,
              sort = list(everything() ~ "frequency")) %>%
  bold_labels() %>%
  add_stat_label() %>%
  modify_header(label ~ "**Cancer characteristics**",
                all_stat_cols() ~ "**{level}**, N = {n}") %>% 
  modify_spanning_header(everything() ~ "Primary Diagnosis Site")
```
<br>

## Let's now claen for histology behavior
```{r clean histology, echo = TRUE, class.source= 'fold-hide'}
gyn_data <- gyn_data %>%
  filter(histology_behavior %in% c(
    "84413 Serous cystadenocarcinoma, NOS", "83803 Endometrioid adenocarcinoma, NOS",
    "81403 Adenocarcinoma, NOS", "83103 Clear cell adenocarcinoma, NOS",
    "84413 Serous carcinoma, NOS", "84603 Papillary serous cystadenocarcinoma",
    "84613 Serous surface papillary carcinoma", "80103 Carcinoma, NOS",
    "89803 Carcinosarcoma, NOS", "83233 Mixed cell adenocarcinoma",
    "84603 Low grade serous carcinoma", "84803 Mucinous adenocarcinoma",
    "84703 Mucinous cystadenocarcinoma, NOS", "89503 Mullerian mixed tumor",
    "82603 Papillary adenocarcinoma, NOS", "80703 Squamous cell carcinoma, NOS",
    "81203 Transitional cell carcinoma, NOS", "84503 Papillary cystadenocarcinoma, NOS",
    "80213 Carcinoma, anaplastic, NOS", "84613 High grade serous carcinoma",
    "84806 Pseudomyxoma peritonei"
  ))



```
We now have `r nrow(gyn_data)` patients after filtering for histology behavior.

```{r}
gyn_data %>% 
  select(disease_type, tumor_specimen_site_of_collection#, histology_behavior
         ) %>% 
  tbl_summary(sort = list(everything() ~ "frequency")) %>%
  bold_labels() %>%
  add_stat_label() %>%
  modify_header(label ~ "**Disease characteristics**",
                all_stat_cols() ~ "**{level}**, N = {n}")
```
<br>

```{r}
gyn_data %>% 
  # mutate(primary_diagnosis_site = relevel(factor(primary_diagnosis_site), ref = "Ovary")) %>% 
  mutate(primary_diagnosis_site = forcats::fct_infreq(primary_diagnosis_site)) %>% 
  # mutate(primary_diagnosis_site = relevel(factor(primary_diagnosis_site), ref = "Ovary")) %>% 
  select(disease_type, 
         histology, 
         primary_diagnosis_site, 
         
         has_subsequent_cancer, subsequent_cancer_primary_diagnosis_site, 
         
         primary_diagnosis_site_code,
         clin_group_stage, other_staging_system, other_staging_value, 
         ecog, karnofsky) %>% 
  tbl_summary(by = primary_diagnosis_site,
              sort = list(everything() ~ "frequency")) %>%
  bold_labels() %>%
  add_stat_label() %>%
  modify_header(label ~ "**Cancer characteristics**",
                all_stat_cols() ~ "**{level}**, N = {n}") %>% 
  modify_spanning_header(everything() ~ "Primary Diagnosis Site")
```
<br>


# Patient characteristics
I have age at medication and age at surgery-biopsy so I should be able to categorize Neoadjuvant and Adjuvant therapies.
```{r}
gyn_data %>% 
  select(age_at_diagnosis, sex, race, ethnicity,
         bmi, bmi_cat,
         smoking_status,
         vital_status,
         
         drugs_ever, age_at_med_start_1,
         surgery_ever, age_at_surgery_biopsy,
         os_event, os_time_from_treatment_years, 
         pfs_event, pfs_time_from_treatment_years
         ) %>% 
  tbl_summary(
              sort = list(all_categorical() ~ "frequency")) %>%
  bold_labels() %>%
  add_stat_label() %>%
  modify_header(label ~ "**Patient characteristics**",
                all_stat_cols() ~ "**{level}**, N = {n}")
```
<br>

# Tumor markers
I need to look more at these markers
```{r}
gyn_data %>% 
  select(fish_t_4_14 : hpv_63) %>% 
  purrr::keep(~!all(is.na(.))) %>%
  tbl_summary() %>%
  bold_labels() %>%
  # add_stat_label() %>%
  modify_header(label ~ "**Tumor markers and cytogenetics**",
                all_stat_cols() ~ "**{level}**, N = {n}")
```

# Medications before germline
```{r}
gyn_data %>% 
  select(drugs_ever) %>% 
  tbl_summary() %>%
  bold_labels()

gyn_data %>% 
  filter(drugs_ever == "Yes") %>% 
  select(avatar_key, starts_with("regimen")) %>% 
  pivot_longer(cols = -avatar_key,
               values_to = "drug_name",
               values_drop_na = TRUE) %>% 
  group_by(avatar_key) %>% 
  mutate(number_of_regimen = n()) %>% 
  ungroup() %>% 
  distinct(avatar_key, .keep_all = TRUE) %>% 
  select(number_of_regimen) %>% 
  tbl_summary(type = list(number_of_regimen ~ "categorical")) %>%
  bold_labels()

gyn_data %>% 
  filter(drugs_ever == "Yes") %>% 
  select(avatar_key, starts_with("regimen")) %>% 
  pivot_longer(cols = -avatar_key, 
               values_to = "drug_name",
               values_drop_na = TRUE) %>% 
  select(drug_name) %>% 
  tbl_summary(sort = list(everything() ~ "frequency")) %>%
  bold_labels()

gyn_data %>% 
  filter(drugs_ever == "Yes") %>% 
  select(avatar_key, starts_with("regimen")) %>% 
  pivot_longer(cols = -avatar_key, 
               values_to = "drug_name",
               values_drop_na = TRUE) %>% 
  mutate(received_carboplatin = case_when(
    str_detect(drug_name, "Carboplatin")        ~ "Yes"
  )) %>% 
  group_by(avatar_key) %>% 
  fill(received_carboplatin, .direction = "updown") %>% 
  ungroup() %>% 
  distinct(avatar_key, .keep_all = TRUE) %>% 
  mutate(received_carboplatin = str_replace_na(received_carboplatin, "No")) %>% 
  select(received_carboplatin) %>% 
  tbl_summary() %>%
  bold_labels()
```





