---
title: "Investigate-patients-with-ovarian-samples"
author: "Christelle Colin-Leitzinger"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: paged
    code_folding: hide
  word_document:
    toc: true
editor_options:
  chunk_output_type: console
---

<style type= "text/css">

. figure {
   margin-top: 25px;
   margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```

```{r library}
library(here)
library(tidyverse)
library(gtsummary)
library(labelled)
library(survival)
library(survminer)
theme_set(theme_classic())
theme_gtsummary_compact()
```

```{r load}
gyn_data <-
  read_rds(paste0(here(), 
                  "/data/processed data",
                  "/ORIEN_data_with_ovarian_sample2025-07-08.rds")) %>% 
  janitor::clean_names() %>% 
  mutate(ch_status = factor(ch_status, levels = c("No CH", "CH")))

path_raw <- fs::path("", "Volumes", "Lab_Gillis",
                     "ORIEN_Avatar", "resources")
load("~/Documents/GitHub/Gillis/ch_orien/Rmd/R01 Gillis-Teng 2023-2024/initial_files.RData")
```

We have `r nrow(gyn_data)` patients with germline and tumor samples coolected in "GYN - Ovarian Cancer" disease type.



# Disease and histology
These are characteristics listed in the sequencing sample data.    
Some samples are not collected in the ovary. I talked with Yi-Han then Nancy. Yi-Han updated her samples selection filters to include ovary samples as first choice when available.
```{r}
gyn_data %>% 
  select(disease_type, tumor_specimen_site_of_collection, histology_behavior
         ) %>% 
  tbl_summary(sort = list(everything() ~ "frequency")) %>%
  bold_labels() %>%
  add_stat_label() %>%
  modify_header(label ~ "**Disease characteristics**",
                all_stat_cols() ~ "**{level}**, N = {n}")
```
<br>

# Let's clean primary cancer first
## Cancer characteristics before primary cancer filtering
You can see that 18 patients had a primary diagnosis of breast cancer and 12 of them had a subsequent diagnosis of ovarian cancer.    
I can extract the histology of the subsequent diagnosis of ovarian cancer if we want to include them.
```{r}
gyn_data %>% 
  # mutate(primary_diagnosis_site = relevel(factor(primary_diagnosis_site), ref = "Ovary")) %>% 
  mutate(primary_diagnosis_site = forcats::fct_infreq(primary_diagnosis_site)) %>% 
  # mutate(primary_diagnosis_site = relevel(factor(primary_diagnosis_site), ref = "Ovary")) %>% 
  select(disease_type, 
         histology, 
         primary_diagnosis_site, 
         
         has_subsequent_cancer, subsequent_cancer_primary_diagnosis_site, 
         
         primary_diagnosis_site_code) %>% 
  tbl_summary(by = primary_diagnosis_site,
              sort = list(everything() ~ "frequency")) %>%
  bold_labels() %>%
  add_stat_label() %>%
  modify_header(label ~ "**Cancer characteristics**",
                all_stat_cols() ~ "**{level}**, N = {n}") %>% 
  modify_spanning_header(everything() ~ "Primary Diagnosis Site")
```
<br>

```{r clean primary site, echo = TRUE, class.source= 'fold-show'}
gyn_data <- gyn_data %>%
  filter(primary_diagnosis_site %in% c(
    "Ovary", "Fallopian tube", "Peritoneum, NOS", "Specified parts of peritoneum"
  ))
```

## Cancer characteristics after primary cancer filtering

We now have `r nrow(gyn_data)` patients after filtering for Primary diagnosis site.

```{r}
gyn_data %>% 
  select(disease_type, tumor_specimen_site_of_collection#, histology_behavior
         ) %>% 
  tbl_summary(sort = list(everything() ~ "frequency")) %>%
  bold_labels() %>%
  add_stat_label() %>%
  modify_header(label ~ "**Disease characteristics**",
                all_stat_cols() ~ "**{level}**, N = {n}")
```
<br>

```{r}
gyn_data %>% 
  # mutate(primary_diagnosis_site = relevel(factor(primary_diagnosis_site), ref = "Ovary")) %>% 
  mutate(primary_diagnosis_site = forcats::fct_infreq(primary_diagnosis_site)) %>% 
  # mutate(primary_diagnosis_site = relevel(factor(primary_diagnosis_site), ref = "Ovary")) %>% 
  select(disease_type, 
         histology, 
         primary_diagnosis_site, 
         
         has_subsequent_cancer, subsequent_cancer_primary_diagnosis_site, 
         
         primary_diagnosis_site_code) %>% 
  tbl_summary(by = primary_diagnosis_site,
              sort = list(everything() ~ "frequency")) %>%
  bold_labels() %>%
  add_stat_label() %>%
  modify_header(label ~ "**Cancer characteristics**",
                all_stat_cols() ~ "**{level}**, N = {n}") %>% 
  modify_spanning_header(everything() ~ "Primary Diagnosis Site")
```
<br>

## Let's now clean for histology behavior

What is the good histology to use?    
histology_behavior is from the sequencing data and histology is from the Diagnosis file which I guess is more clinical.
```{r}
gyn_data %>%
  mutate(histology_behavior_sequencing = str_match(histology_behavior, "([:digit:]*) (.*)")[,3]) %>% 
  mutate(diff = case_when(
    histology_behavior_sequencing == histology       ~ "Same",
    TRUE                                             ~ "Diff"
  )) %>% 
  filter(diff == "Diff") %>% 
  select(histology_behavior_sequencing, histology)
  
# str_match(gyn_data$histology_behavior, "([:digit:]*) (.*)")
```


```{r clean histology, echo = TRUE, class.source= 'fold-show'}
gyn_data <- gyn_data %>%
  filter(histology_behavior %in% c(
    "84413 Serous cystadenocarcinoma, NOS", "83803 Endometrioid adenocarcinoma, NOS",
    "81403 Adenocarcinoma, NOS", "83103 Clear cell adenocarcinoma, NOS",
    "84413 Serous carcinoma, NOS", "84603 Papillary serous cystadenocarcinoma",
    "84613 Serous surface papillary carcinoma", "80103 Carcinoma, NOS",
    "89803 Carcinosarcoma, NOS", "83233 Mixed cell adenocarcinoma",
    "84603 Low grade serous carcinoma", "84803 Mucinous adenocarcinoma",
    "84703 Mucinous cystadenocarcinoma, NOS", "89503 Mullerian mixed tumor",
    "82603 Papillary adenocarcinoma, NOS", "80703 Squamous cell carcinoma, NOS",
    "81203 Transitional cell carcinoma, NOS", "84503 Papillary cystadenocarcinoma, NOS",
    "80213 Carcinoma, anaplastic, NOS", "84613 High grade serous carcinoma",
    "84806 Pseudomyxoma peritonei"
  ))



```
We now have `r nrow(gyn_data)` patients after filtering for histology behavior.

```{r}
gyn_data %>% 
  select(disease_type, tumor_specimen_site_of_collection#, histology_behavior
         ) %>% 
  tbl_summary(sort = list(everything() ~ "frequency")) %>%
  bold_labels() %>%
  add_stat_label() %>%
  modify_header(label ~ "**Disease characteristics**",
                all_stat_cols() ~ "**{level}**, N = {n}")
```
<br>

```{r}
gyn_data %>% 
  # mutate(primary_diagnosis_site = relevel(factor(primary_diagnosis_site), ref = "Ovary")) %>% 
  mutate(primary_diagnosis_site = forcats::fct_infreq(primary_diagnosis_site)) %>% 
  # mutate(primary_diagnosis_site = relevel(factor(primary_diagnosis_site), ref = "Ovary")) %>% 
  select(disease_type, 
         histology, 
         primary_diagnosis_site, 
         
         has_subsequent_cancer, subsequent_cancer_primary_diagnosis_site, 
         
         primary_diagnosis_site_code) %>% 
  tbl_summary(by = primary_diagnosis_site,
              sort = list(everything() ~ "frequency")) %>%
  bold_labels() %>%
  add_stat_label() %>%
  modify_header(label ~ "**Cancer characteristics**",
                all_stat_cols() ~ "**{level}**, N = {n}") %>% 
  modify_spanning_header(everything() ~ "Primary Diagnosis Site")
```


<br>


# Patient characteristics
BMI is selected to be the closest to Dx (before or after). 
But in case of ovarian cancer for which patient often receive surgery, should I take only BMI before Dx.
Or make a rule depending if patient received surgery?
```{r}
gyn_data %>% 
  select(age_at_diagnosis, sex, race, ethnicity,
         race_eth,
         clin_group_stage, other_staging_system, other_staging_value, 
         ecog, karnofsky,
         bmi, bmi_cat,
         smoking_status,
         vital_status,
         not_real_age, #not_real_sequencing_age,
         drugs_ever, age_at_med_start_1,
         surgery_ever, age_at_first_surgerybiopsy,
         time_dx_to_first_treatment,
         os_event, os_time_from_treatment_years, 
         pfs_event, pfs_time_from_treatment_years,
         ch_status, tumor_ch_status
         ) %>% 
  tbl_summary(by = ch_status,
              sort = list(all_categorical() ~ "frequency")) %>%
  bold_labels() %>%
  add_p() %>% bold_p(t=0.05) %>% add_overall() %>% 
  # add_stat_label() %>%
  modify_header(label ~ "**Patient characteristics**",
                all_stat_cols() ~ "**{level}**, N = {n}")
```
<br>

# Treatment
Should we make a rule to exclude treatment that would have happen before ovarian cancer diagnosis?

## Surgery
I linked the surgery to the Ovary diagnosis.    
FYI there are some SiteTherapeutic == "Unknown/Not Applicable". For these we used to code surgery ever "Yes" if we have a age related to this surgery and "No" if there was no age associated     
Also there is some surgery/biopy with PrimaryDiagnosisSite == "Unknown/Not Applicable" so i cannot merge them to a specific cancer diagnosis.
The choice we made a long time ago was to take the first surgery with the correct PrimaryDiagnosisSite and if none to select the first surgery with unknown primary site.
We could also create a rule for a specific location in case of unknown PrimaryDiagnosisSite.    
For surgery, there is variables called SiteDiagnostic, SiteTherapeutic, SitePalliative. 
I think this is where we can understand if "Yes" surgery-biopsy is an actual surgery or a biopsy. 
What do you think? Should we add the SiteTherapeutic == "Yes" filter?     


```{r}
SurgeryBiopsy %>%
  # For the patients with no diagnosis site
  filter(PrimaryDiagnosisSite == "Unknown/Not Applicable") %>% 
  select(SurgeryBiopsyLocation, SiteDiagnostic, surgery_ever = SiteTherapeutic, SitePalliative,
         MethodCoreBiopsyTissue, MethodCoreBiopsyLymphNode, MethodSurgicalResection,
         MethodSentinelLymphNode, MethodSurgeryNOS, MethodOther, )
```
This is what I have for now but this will change depending of the new cleaning steps.
```{r}
gyn_data %>% 
  select(age_at_first_surgerybiopsy, has_surgery_data,
         surgerybiopsy_ever, 
         surg_primary_diagnosis_site,
         surgery_ever,
         ch_status) %>%
  tbl_summary(by = ch_status) %>%
  bold_labels() %>%
  add_p() %>% bold_p(t = 0.05) %>% add_overall() %>% 
  modify_header(label ~ "**Surgery**",
                all_stat_cols() ~ "**{level}**, N = {n}")
```

## Medication
Do we care about the mediacation which was received before germline?    
We didn't linked the mediacation to the Ovary diagnosis for these (probably think that most drugs have an antineoplastic effect on the whole physiologie.    
FYI there are also some SiteTherapeutic == "Unknown/Not Applicable". Previously we included all the medication received before germline.   
I recreated the regimen by same start age.   
```{r}
gyn_data %>% 
  select(age_at_med_start_1, has_medication_data,
         drugs_ever, 
         regimen_1, ch_status) %>%
  tbl_summary(by = ch_status) %>%
  bold_labels() %>%
  add_p() %>% bold_p(t = 0.05) %>% add_overall() %>% 
  modify_header(label ~ "**Medication**",
                all_stat_cols() ~ "**{level}**, N = {n}")
```

### Medications after germline
```{r}
gyn_data %>% 
  filter(drugs_ever == "Yes") %>% 
  select(avatar_key, starts_with("regimen")) %>% 
  pivot_longer(cols = -avatar_key,
               values_to = "drug_name",
               values_drop_na = TRUE) %>% 
  group_by(avatar_key) %>% 
  mutate(number_of_regimen = n()) %>% 
  ungroup() %>% 
  distinct(avatar_key, .keep_all = TRUE) %>% 
  select(number_of_regimen) %>% 
  tbl_summary(type = list(number_of_regimen ~ "categorical")) %>%
  bold_labels()

gyn_data %>% 
  filter(drugs_ever == "Yes") %>% 
  select(avatar_key, starts_with("regimen")) %>% 
  pivot_longer(cols = -avatar_key, 
               values_to = "drug_name",
               values_drop_na = TRUE) %>% 
  select(drug_name) %>% 
  tbl_summary(sort = list(everything() ~ "frequency")) %>%
  bold_labels()

gyn_data %>% 
  filter(drugs_ever == "Yes") %>% 
  select(avatar_key, starts_with("regimen"), ch_status) %>% 
  pivot_longer(cols = -c(avatar_key, ch_status), 
               values_to = "drug_name",
               values_drop_na = TRUE) %>% 
  mutate(received_carboplatin = case_when(
    str_detect(drug_name, "Carboplatin")        ~ "Yes"
  )) %>% 
  group_by(avatar_key) %>% 
  fill(received_carboplatin, .direction = "updown") %>% 
  ungroup() %>% 
  distinct(avatar_key, .keep_all = TRUE) %>% 
  mutate(received_carboplatin = str_replace_na(received_carboplatin, "No")) %>% 
  select(received_carboplatin, ch_status) %>% 
  tbl_summary(by = ch_status) %>%
  add_p() %>% bold_p(t = 0.05) %>% add_overall()
```





## Radiation
I don't think we are interested by radiation?
```{r}
gyn_data %>% 
  select(age_at_first_radiation, has_radiation_data,
         radiation_ever,
         ch_status) %>%
  tbl_summary(by = ch_status) %>%
  bold_labels() %>%
  add_p() %>% bold_p(t = 0.05) %>% add_overall() %>% 
  modify_header(label ~ "**Radiation**",
                all_stat_cols() ~ "**{level}**, N = {n}")
```

## SCT
No SCT
```{r}
gyn_data %>% 
  select(age_at_first_sct, has_sct_data,
         sct_ever,
         ch_status) %>%
  tbl_summary(by = ch_status) %>%
  bold_labels() %>%
  add_p() %>% bold_p(t = 0.05) %>% add_overall() %>% 
  modify_header(label ~ "**SCT**",
                all_stat_cols() ~ "**{level}**, N = {n}")
```


# Tumor markers
I need to look more at these markers
```{r}
gyn_data %>% 
  select(fish_t_4_14 : hpv_63) %>% 
  purrr::keep(~!all(is.na(.))) %>%
  tbl_summary() %>%
  bold_labels() %>%
  # add_stat_label() %>%
  modify_header(label ~ "**Tumor markers and germline_collection_age**",
                all_stat_cols() ~ "**{level}**, N = {n}")
```

# Survival

For PFS I have "Progression" and "Recurrence", both are coded as event = 1.   
I linked the outcomes info by cancer diagnosis so I should have only the progression/recurrence for ovarian cancer.   
FYI, some are with OutcomesPrimaryDiagnosisSite == "Unknown/Not Applicable" so i cannot merge them to a specific cancer diagnosis.
Same choice was made as for surgery: select the info with correct PrimaryDiagnosisSite and if none select the first outcome with unknown primary site.
 
```{r}
gyn_data %>% select(pfs_time_from_treatment_years) %>% arrange(pfs_time_from_treatment_years)
gyn_data %>% select(os_time_from_treatment_years) %>% arrange(os_time_from_treatment_years)
gyn_data %>% 
  select(vital_status,
         has_os_data,
         os_event, os_time_from_treatment_years, 
         has_outcomes_data, 
         pfs_event, pfs_time_from_treatment_years,
         ch_status, tumor_ch_status
         ) %>% 
  tbl_summary(by = ch_status,
              sort = list(all_categorical() ~ "frequency")) %>%
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05) %>% add_overall() %>% 
  modify_header(label ~ "**Survival**")
```

## HR - Univariable only - need better cleaning of the factors included for multivariable analysis
```{r}
tbl_uvregression(
  gyn_data,
  method = coxph,
  y = Surv(time = os_time_from_treatment_years, event = os_event),
  exponentiate = TRUE,
  include = c("ch_status", "germline_collection_age", "race_eth",
              "clin_group_stage", "histology",
              "bmi_cat"),
  # pvalue_fun = label_style_pvalue(digits = 2)
  ) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ "**OS**")

# coxph(Surv(time = gyn_data$os_time_from_treatment_years, 
#                    event = gyn_data$os_event) ~ ch_status + germline_collection_age + race_eth +
#         clin_group_stage + histology + bmi_cat, 
#               data = gyn_data)  %>% 
#   tbl_regression(exponentiate = TRUE) %>% 
#   bold_p(t = .05) %>% 
#   add_nevent(location = "level") %>% add_n(location = "level") %>% 
#   modify_spanning_header(everything() ~ "**OS**")

tbl_uvregression(
  gyn_data,
  method = coxph,
  y = Surv(time = pfs_time_from_treatment_years, event = pfs_event),
  exponentiate = TRUE,
  include = c("ch_status", "germline_collection_age", "race_eth",
              "clin_group_stage", "histology",
              "bmi_cat"),
  # pvalue_fun = label_style_pvalue(digits = 2)
  ) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ "**PFS**")

# coxph(Surv(time = gyn_data$pfs_time_from_treatment_years, 
#                    event = gyn_data$pfs_event) ~ ch_status + germline_collection_age + race_eth +
#         clin_group_stage + histology + bmi_cat, 
#               data = gyn_data)  %>% 
#   tbl_regression(exponentiate = TRUE) %>% 
#   bold_p(t = .05) %>% 
#   add_nevent(location = "level") %>% add_n(location = "level") %>% 
#   modify_spanning_header(everything() ~ "**PFS**")
```

## KM
```{r}
ggsurvplot(survfit(Surv(os_time_from_treatment_years, os_event) ~ ch_status, 
                   data = gyn_data),
           # title = "PFS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(15, "bold", "black"),
           font.y = c(15, "bold", "black"),
           font.legend = c(14, "black"),
           font.tickslab = c(12, "bold", "black"),
           size = 1.5,

           xlab = "Time (years)",
           ylab = "OS (probability)",
           legend = "top",
           # legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) %++% guides(colour = guide_legend(nrow = 1))

ggsurvplot(survfit(Surv(pfs_time_from_treatment_years, pfs_event) ~ ch_status, 
                   data = gyn_data),
           # title = "PFS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(15, "bold", "black"),
           font.y = c(15, "bold", "black"),
           font.legend = c(14, "black"),
           font.tickslab = c(12, "bold", "black"),
           size = 1.5,

           xlab = "Time (years)",
           ylab = "PSS (probability)",
           legend = "top",
           # legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) %++% guides(colour = guide_legend(nrow = 1))
```



